### 2015-07-09 narrative/data-description.tex
minor changes

From an application level, we will explore these techniques through our ADIOS
framework using many of our existing application partners, including
XGC1~\cite{chang2006integrated}, GTC~\cite{klasky2003grid} , and SPECFM3D~\cite{SPECFEM3D}. We will investigate: 1) what hints the users can give
the system, and 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
tier and lifetime, which can be specified by the user through the API.
As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
The user can pick a data-refactoring scheme (see
Sec~\ref{sec:data-refactor}) which will classify the particle data into four different
prioritization levels. The user can then pick a data utility, such as three months
for the highest priority items and 12 months for the lowest priority data.
The user will then call an API function that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return a time estimate for this write operation.
If this time estimate is too large, an API call can be made to 
specify the lowest priority element to be written, skipping the rest in order to save time.
The data will then be stored
across the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
SSIO will decide if the data has to go through the parallel file system 
towards a lower stack or can be staged to that stack directly. 
% Description of the utility function research

@@

The assumptions which we will make in choosing a utility function is that:
1) There will only be a small number of priority levels that users will use
to re-prioritize their data; 2) The number of tiers in the storage system will
also be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
lifetime of objects in each tier will be specified by the resource by
a tuple (size, lifetime).  The user can then utilize this information to
place information to the system such as: store the highest priority data
as high up on the hierarchy for 6 months, and then migrate this data to tape.
The user can also define the utility for CR files as that they have high
priority but their lifetime is for a short period of time; i.e. when the
next CR file appears on the system or the next two CR files. Research in
describing this utility function will be in conjunction with our users
and we will evaluate how policies can be placed in the system and how well
this utility function can be observed by the user.

>>>>

From an application level, we will explore these techniques through our ADIOS
framework using many of our existing application partners, including
XGC1~\cite{chang2006integrated}, GTC~\cite{klasky2003grid} , and SPECFM3D~\cite{SPECFEM3D}. We will investigate: 1) what hints the users can give
the system, and 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
tier and lifetime, which can be specified by the user through the API.
As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
The user can pick a data-refactoring scheme (see
\S~\ref{sec:data-refactor}) which will classify the particle data into four different
prioritization levels. The user can then pick a data utility, such as three months
for the highest priority items and 12 months for the lowest priority data.
The user will then call an API function that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return a time estimate for this write operation.
If this time estimate is too large, an API call can be made to 
specify the lowest priority element to be written, skipping the rest in order to save time.
The data will then be stored
across the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
SSIO will decide if the data has to go through the parallel file system 
towards a lower stack or can be staged to that stack directly. 
% Description of the utility function research

@@

The assumptions which we will make in choosing a utility function is that:
1) There will only be a small number of priority levels that users will use
to re-prioritize their data; 2) The number of tiers in the storage system will
be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
lifetime of objects in each tier will be specified by the resource by
a tuple (size, lifetime).  The user can then utilize this information to
place information to the system such as: store the highest priority data
as high up on the hierarchy for 6 months, and then migrate this data to tape.
The user can also define the utility for CR files as that they have high
priority but their lifetime is for a short period of time; i.e. when the
next CR file appears on the system or the next two CR files. Research in
describing this utility function will be in conjunction with users
and we will evaluate how policies can be placed in the system and how well
this utility function can be observed by the user.

### 2015-07-09 narrative/intro.tex
minor changes

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\caption {Relevance to SSIO}
\label{tab:call}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of I/O middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\
\hline
\end{tabular}
\end{center}
\vskip -0.5cm
\end{table}

@@

The latest use case is from  the XGC1~\cite{chang2006integrated} simulation from C. S. Chang (PPPL), who is one of the largest users of Leadership Class Facilities (over 300M hours at DOE facilities). They launched a series of simulations, planning to produce 100~PB of data over 10 days of total runtime on the Titan system. An expert team was assembled from the user group, the  I/O, and storage  team to help with ensuring that the maximum amount of information would be
saved. Due to physical resource limitations, the simulation was restricted to write about 10~PB initially. However, once the time to write and read, together with the financial cost to archive this much data were fully explored, this plan was significantly reduced. At each stage, the particular expert knowledge regarding the file system and the archival storage were not known to the user team. As these particulars were realized, significant modifications in the plans were made. Eventually, they were forced to make very hard decisions in reducing the desired 100~PB over a 10-day run to only 5~PB of data.

>>>>

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\caption {Relevance to SSIO}
\label{tab:call}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant SIRIUS work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of middleware with the storage system & Manage data from the application  to middleware to the storage system using an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & Define a utility functions for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\
\hline
\end{tabular}
\end{center}
\vskip -0.5cm
\end{table}

@@

The latest use case is from  the XGC1~\cite{chang2006integrated} simulation from C. S. Chang (PPPL), who is one of the largest users of Leadership Class Facilities (over 300M hours at DOE facilities). They launched a series of simulations, planning to produce 100~PB of data over 10 days of total runtime on the Titan system. An expert team was assembled from the user group, the  I/O, and storage  team to help with ensuring that the maximum amount of information would be
saved. Due to physical resource limitations, the simulation was restricted to write about 10~PB initially. However, once the time to write and read, together with the financial cost to archive this much data were fully explored, this plan was significantly reduced. At each stage, the particular expert knowledge regarding the file system and the archival storage were not known to the user team. As these particulars were realized, significant modifications in the plans were made. Eventually, they were forced to make very hard decisions in reducing the desired 100~PB to only 5~PB of data.
